Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-14 ieee_paper.tex
4.a

In this work, we use a pre-release KNC card SE10P. The card has 8
memory controllers where each of them can execute 5.5 billion
transactions per second and has two 32-bit channels. That is the
architecture can achieve a total bandwidth of 352GB/s aggregated
across all the memory controllers. There are 61 cores clocked at
1.05GHz. The cores’ memory interface are 32-bit wide with two channels
and the total bandwidth is 8.4GB/s per core. Thus, the cores should be
able to consume 512.4GB/s at most. However, the bandwidth between
the cores and the memory controllers is limited by the ring network
which connects them and theoretically supports at most 220GB/s.

Each core in the architecture has a 32kB L1 data cache, a 32kB L1
instruction cache, and a 512kB L2 cache. The architecture of a core is
based on the Pentium architecture: though its design has been updated
to 64-bit. A core can hold 4 hardware contexts at any time. And at
each clock cycle, instructions from a single thread are executed. Due
to the hardware constraints and to overlap latency, a core never
executes two instructions from the same hardware context
consecutively. In other words, if a program only uses one thread, half
of the clock cycles are wasted. Since there are 4 hardware contexts
available, the instructions from a single thread are executed
in-order. As in the Pentium architecture, a core has two di↵erent
concur- rent instruction pipelines (called U-pipe and V-pipe) which
allow the execution of two instructions per cycle. However, some
instructions are not available on both pipelines: only one vector or
floating point instruction can be executed at each cycle, but two ALU
instructions can be executed in the same cycle.

Most of the performance of the architecture comes from the vector
processing unit. Each of Intel Xeon Phi’s cores has 32 512-bit SIMD
registers which can be used for double or single precision, that is,
either as a vector of 8 64-bit values or as a vector of 16 32-bit
values, respectively. The vector processing unit can perform many
basic instructions, such as addition or division, and mathematical
operations, such as sine and sqrt, allowing to reach 8 double
precision operations per cycle (16 single precision). The unit also
sup- ports Fused Multiply-Add (FMA) operations which are typically
counted as two operations for benchmarking purposes. Therefore, the
peak performance of the SE10P card is 1.0248 Tflop/s in double
precision (2.0496 Tflop/s in single precision) and half without FMA.

>>>>

In this work, we use an Intel Xeon Phi 5110P coprocessor. This card
has 8 memory controllers where each of them can execute 5 billion
transactions per second and has two 32-bit channels, achieving a total
bandwidth of 320GB/s aggregated across all the memory
controllers. There are 60 cores clocked at 1.05GHz. Their memory
interfaces are 32-bit wide with two channels and the total bandwidth
is 8.4GB/s per core. Thus, the cores should be able to consume 504GB/s
at most. However, the bandwidth between the cores and the memory
controllers is limited by the ring network that connect the cores and
the memory controller. Its precise bandwidth is unknown for this
precise card, but it would be coehrent that it is between 200GB/s and
250GB/s.

Each core in the architecture has a 32kB L1 data cache, a 32kB L1
instruction cache, and a 512kB L2 cache. The architecture of a core is
based the traditional Pentium architecture which have been extended to
64-bit. A core can hold 4 hardware contexts at any time. And at each
clock cycle, instructions from a single thread are executed. Due to
some hardware constraints two hardware context must be used to reach
the peak instruction throughput in that architecture. As in the
Pentium architecture, a core has two different concurrent instruction
pipelines which allow the execution of two instructions per
cycle. However, only one vector or floating point instruction can be
executed at each cycle.

Most of the performance of the architecture comes from the vector
processing unit. Each core has 32 512-bit SIMD registers which can be
used for double or single precision, that is, either as a vector of 8
64-bit values or as a vector of 16 32-bit values, respectively. The
vector processing unit can perform many basic instructions, such as
addition or division, and mathematical operations, such as sine and
sqrt, allowing to reach 8 double precision operations per cycle (16
single precision). The unit also supports Fused Multiply-Add (FMA)
operations which are typically counted as two operations for
benchmarking purposes. Therefore, the peak performance of the 5110P
card is 1 Tflop/s in double precision and 2 Tflop/s in single
precision. If FMA can not be used and half of these number can be
achieved.
